### 2015-03-15 staticanalysis.tex
updates to instrumentation text

To be sound, \whoop returns a nondeterministic value (using the Boogie keyword \texttt{havoc}) for each read access to a shared memory location. This over-approximates any effects from all the unmodeled threads on the shared state, but can and will result into false positives.

%For each array, two Booleans are introduced to record whether a read from or write to the watched offset has occurred. Initially these Booleans are false, and they are reset at each barrier. Thread s sets the ?read? Boolean to true whenever it reads from the watched offset, and similarly for the ?write? Boolean. A race between s and t is reported if thread t reads from the watched offset and the ?write? Boolean is true, or if thread t writes to the watched offset and either the ?read? or ?write? Boolean is true. The non-deterministic choice per array access is eliminated.

%\begin{boogie}[caption = Translation to Boogie IVL using SMACK, label = fig:smack_translation]
%call mutex_lock($p3);
%\end{boogie}

Our verification technique begins by performing \emph{two-thread reduction}, a sound abstraction that removes all but two arbitrary threads, each running an entry point of the originally concurrent program, and then performs \emph{pairwise sequentialisation}, which combines the two arbitrary threads in a single sequential pair. This is achieved by replacing the \texttt{return} statement of the first entry point with a \texttt{goto} statement that jumps execution to the first statement of the second entry point. This process repeats until all pairs of entry points have been sequentialised. To achieve soundness, each time an entry point performs a read access to a shared resource, we return a nondeterministic value. This over-approximates any effects from all the unmodeled threads on the device driver shared state.

Next, we split each pair into two distinct analysis regions: the \emph{logger}, which contains all the statements from the first entry point; and the \emph{checker}, which contains all the statements from the second entry point. The logger and the checker are then instrumented according to Table~\ref{tab:instrumentation}. The two regions directly contribute to the static lockset analysis and race checking as follows: the logger is responsible for recording all locksets $\mathit{LS}$ maintained by its entry point; while the checker asserts for potential conflicts between its own entry point and a given lockset $\mathit{LS}$ from the logger. This whole process is summarised below:

\begin{description}
  \item[Logger] Initially, for each memory location $m$ accessed by the logger's entry point, the logger's lockset $\mathit{LS}\lbrack m\rbrack$ contains all locks $\ell$ in the device driver, while the write access set $W\lbrack m\rbrack$ is set to false. Also, the logger's current lockset $\mathit{LS}_{CR}$, which is the set of all locks \emph{currently} held by the logger's entry point, is initially empty. Each time the logger performs a read or write access to $m$, it updates its lockset $\mathit{LS}\lbrack m\rbrack$, with the intersection of $\mathit{LS}\lbrack m\rbrack$ and $\mathit{LS}_{CR}$. This process is known as \emph{lockset refinement}~\cite{savage1997eraser} and results into a $\mathit{LS}\lbrack m\rbrack$ that contains only the locks $\ell$ that consistently protect $m$ during the execution of the logger's entry point.

  \item[Checker] Initially, for each memory location $m$ accessed by the checker's entry point, the checker's lockset $\mathit{LS}\lbrack m\rbrack$ and write access set $W\lbrack m\rbrack$ are equal to the corresponding sets supplied at the end of the logger. Also, the logger's current lockset $\mathit{LS}_{CR}$ is initially empty. Each time the checker performs a write access to $m$, it asserts that the intersection between $\mathit{LS}_{CR}$ and $\mathit{LS}\lbrack m\rbrack$ is non-empty. If the assertion fails, it implies that the entry point of the logger and the entry point of the checker potentially race on $m$. In this case a counterexample is generated and reported to the user. To check a read access, the checker performs the aforementioned assertion only if the write access set $W\lbrack m\rbrack$ is true, thus avoids to report a read-read data race, which is inherently benign.
\end{description}

>>>>

\whoop continues by performing \emph{two-thread reduction}, a sound abstraction that removes all but two arbitrary threads, each running an entry point of the originally concurrent program, and then performs \emph{pairwise sequentialisation}, which combines the two arbitrary threads in a single sequential pair. This is achieved by

This process repeats until all pairs of entry points have been sequentialised. To be sound, each time an entry point performs a read access to a shared memory location, \whoop returns a nondeterministic value (using the Boogie keyword \texttt{havoc}). This over-approximates any effects from all the unmodeled threads on the shared state.

Each time the checker performs a write access to $m$, it asserts that the intersection between $\mathit{LS}_{CR}$ and $\mathit{LS}\lbrack m\rbrack$ is non-empty. If the assertion fails, it implies that the entry point of the logger and the entry point of the checker potentially race on $m$. In this case a counterexample is generated and reported to the user. To check a read access, the checker performs the aforementioned assertion only if the write access set $W\lbrack m\rbrack$ is true, thus avoids to report a read-read data race, which is inherently benign.
